Tools and Ideas for the Procedural Generation of Shadow Sculptures

### 2015-12-16 paper/paper.tex
Completed everything before final submition

The innermost loop (over all spheres and cylinders) can be accelerated by use of a space partitioning method. Let us restrict ourselves temporarily to looking for intersections of light paths with spheres. By building a data structure before hand that divides the list of spheres based on the locations of their centers, we can disregard most of them during any one intersection check. This technique was inspired by $k$-d trees, which are space partitioning data structures used in one solution to the nearest neighbor search problem. It turns out that space partitioning methods are very commonly used in ray tracing.

@@

\section{Genetic Algorithm Solutions} \label{GA}
We believe approximate solutions to the forward problem may be possible using genetic algorithms for creating ``abstract'' data that fits a given shadow.  The question then becomes how to represent data for each possible solution to the problem in a genome that allows simple crossovers and mutations.  Some possible representations are as follows:

The simplest method of representing nodes and edges would be to keep the number of both constant, moving the nodes around while retaining the same edge mappings for all genomes.  This way, one could represent each node as an unsigned integer, or an array of three unsigned integers representing the positive incremental distance from some origin point.  Given the nature of the problem, one can disregard any space not within the projection from the light source to a given section of the wall.  This allows the bounding of the entire system in a cube, which we can use to define our \texttt{0} and our \texttt{MAX\_INT} for each coordinate direction.  Next, we would create an array of unsigned integers, of length 3 times the number of nodes in the system.  This would be the genome for the genetic algorithm.  The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the shadow mapping methods above.  Using such a method would prevent situations where the graph became unconnected, but it does not address issues of structural integrity any further than this.  Since the nature of the problem allows for many local solutions, the ability of such a genetic algorithm to converge may be slowed.

One may increase the variability of the genetic algorithm by introducing edge mapping to the genome.  One may represent an edge between node A and node B as a 1 in a symmetric n-by-n adjacency matrix, where n is the number of nodes in the graph.  This allows for easy genetic recombinations, with the ability to vary the number of edges in the graph.  The issue that this does not take care of is the connectivity of the graph.  One would have to either add a penalty to the fitness function which performed a depth-first search on the nodes and penalized unconnected graphs, or alternatively, one could modify the base constructs of the genetic algorithm, making it continue to perform attempts at combining two genomes until it had children that were all fully connected.  For the first option, the question would become how strongly one would decrease the fitness of an unconnected graph, and whether the number of unconnected pieces should increase that penalty.  The second option would allow us to use the same fitness function, but may add extra levels of complexity to the problem due to the required repeated attemps within single generations.

The last important part of the genetic algorithm is a method of performing mutations on genomic data.  Here are some options in consideration:

By simply mutating the data in the genome, one would cause one or multiple nodes to be moved to a new location, possibly on the same plane, but possibly not.  This would be the simplest method of performing a mutation, but depending on how the nodes are laid out, it may mutate a node outside of the acceptable region given for nodes in the given problem.  Whether this is allowed or not depends on the situational requirements given.

@@

  \item Is it a good idea to combine the two cpu-based shadow-rendering techniques? For example, use the second method to draw spheres and the first to draw cylinders?

  \item In the first shadow rendering method, it would be faster to use Bresenham's algorithm to draw an ellipse once we have its parameters, instead of the naive method we use now.
\end{itemize}

\begin{appendices}
\section{Shadow Calculations} \label{app:ellipse_derivation}

>>>>

The innermost loop (over all spheres and cylinders) can be accelerated by use of a space partitioning method. Let us restrict ourselves temporarily to looking for intersections of light paths with spheres. By building a data structure beforehand that divides the list of spheres based on the locations of their centers, we can disregard most of them during any one intersection check. This technique was inspired by $k$-d trees, which are space partitioning data structures used in one solution to the nearest neighbor search problem. It turns out that space partitioning methods are very commonly used in ray tracing.

@@

\section{Genetic Algorithm Solutions} \label{GA}
We believe approximate solutions to the forward problem may be possible using genetic algorithms for creating ``abstract'' data that fits a given shadow.  The question then becomes how to represent data for each possible solution to the problem in a genome that allows simple crossovers and mutations.  Some possible representations are as follows.

The simplest method of representing nodes and edges would be to keep the number of both constant, moving the nodes around while retaining the same edge mappings for all genomes.  This way one could represent each node as an unsigned integer or an array of three unsigned integers representing the positive incremental distance from some origin point.  Given the nature of the problem, one can disregard any space not within the projection from the light source to a given section of the wall.  This allows the bounding of the entire system in a cube, which we can use to define our \texttt{0} and our \texttt{MAX\_INT} for each coordinate direction.  Next we would create an array of unsigned integers, of length 3 times the number of nodes in the system.  This would be the genome for the genetic algorithm.  The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the shadow mapping methods above.  Using such a method would prevent situations where the graph became unconnected, but it does not address issues of structural integrity any further than this.  Since the nature of the problem allows for many local solutions, the ability of such a genetic algorithm to converge may be slowed.

One may increase the variability of the genetic algorithm by introducing edge mapping to the genome.  An edge between node A and node B may be represented as a 1 in a symmetric n-by-n adjacency matrix, where n is the number of nodes in the graph.  This allows for easy genetic recombinations with the ability to vary the number of edges in the graph.  The issue that this does not take care of is the connectivity of the graph.  One would have to either add a penalty to the fitness function which performed a depth-first search on the nodes and penalized unconnected graphs; alternatively, one could modify the base constructs of the genetic algorithm to continue to perform attempts at combining two genomes until it had children that were all fully connected.  For the first option, the question would become how strongly one would decrease the fitness of an unconnected graph and whether the number of unconnected pieces should increase that penalty.  The second option would allow us to use the same fitness function, but may add extra levels of complexity to the problem due to the required repeated attemps within single generations.

The last important part of the genetic algorithm is a method of performing mutations on genomic data.  Here are some options in consideration.

By simply mutating the data in the genome, one would cause one or multiple nodes to be moved to a new location, possibly on the same plane, but possibly not.  This would be the simplest method of performing a mutation, but depending on how the nodes are laid out it may mutate a node outside of the acceptable region given for nodes in the given problem.  Whether this is allowed or not depends on the situational requirements given.

@@

  \item Is it a good idea to combine the two CPU-based shadow-rendering techniques? For example, use the second method to draw spheres and the first to draw cylinders?

  \item In the first shadow rendering method, would it be faster to use Bresenham's algorithm to draw an ellipse once we have its parameters, instead of the naive method we use now?
\end{itemize}

\begin{appendices}
\section{Ellipse Derivation} \label{app:ellipse_derivation}
